Real-Time Scheduling Algorithms and Battery Consumption of Mobile Devices

### 2017-05-03 paper-RMDON.tex
DVS algorithms and results

\section{DVS Algorithms}
For hard real-time systems, there are two kinds of voltage scheduling approaches depending on the voltage scaling granularity:  intra-task DVS (IntraDVS) and inter-task DVS (InterDVS). The intra-task DVS algorithms adjust the voltage within an individual task boundary, while the inter-task DVS algorithms determine the voltage on a task-by-task basis at each scheduling point. The main difference between them is whether the slack times are used for the current task or for the tasks that follow.  InterDVS algorithms distribute the slack times from the current task for the following tasks, while IntraDVS algorithms use the slack times from the current task for the current task itself.

@@

%\section{Target Algorithms}
%
%Table 2 summarizes the DVS algorithms selected for the comparative  study.    Here,  eight  InterDVS  algorithms  are chosen, two [18, 14] of which are based on the RM schedul-
%ing policy, while the other six algorithms [18, 14, 2, 10] are based on the EDF scheduling policy.   For IntraDVS algorithms, two algorithms are selected, one from path-based IntraDVS algorithms [16], and the other from stochastic methods [5]. In these selected DVS algorithms, one or sometimes more than one slack estimation methods explained in the previous section were used.  In lppsEDF and lppsRM which were proposed by Shin et.  al. in  [18], slack time of a task is estimated using the maximum constant speed and Stretching-to-NTA methods.
%
%The ccRM algorithm proposed by Pillai et.   al. [14]  is similar to lppsRM in the sense that it uses both the maximum constant speed and the Stretching-to-NTA methods. However, while lppsRM can adjust the voltage and clock speed only when a single task is active (Figure 1(a)), ccRM extends the stretching to NTA method to the case where multiple tasks are active (Case-II in Figure 1(b)).
%
%Pillai et. al. also  proposed  two  other  DVS  algorithms [14], ccEDF and laEDF, for EDF scheduling policy. These algorithms estimate slack time of a task using the utilization updating method. While ccEDF adjusts the voltage and clock speed based on run-time variation in processor utilization alone, laEDF takes a more aggressive approach by estimating the amount of work required to be completed before NTA. DRA and AGR , which were proposed by Aydin et.  al. in [2],  are two representative DVS algorithms that are based on the priority-based slack stealing method. The DRA algorithm estimates the slack time of a task using the priority-based slack stealing method along with the maximum constant speed and the Stretching-to-NTA methods.  Aydin et. al. also extended the DRA algorithm and proposed another DVS algorithm called AGR for more aggressive slack estimation and voltage/clock scaling. In AGR, in addition to the priority-based slack stealing, more slack times are identified by computing the amount of work required to be completed before NTA (Case-I in Figure 1(b)). lpSHE is  another  DVS  algorithm  which  is  based  on the priority-based slack stealing method [10].  Unlike DRA and AGR, lpSHE extends the priority-based slack stealing method by adding a procedure that estimates the slack time from lower-priority tasks that were completed earlier than expected. DRA, AGR,  and lpSHE algorithms  are  somewhat similar to one another in the sense that all of them use the maximum constant speed in the off-line phase and the Stretching-to-NTA method in the on-line phase in addition to the priority-based slack stealing method. For  IntraDVS  algorithms,  Shin’s  intra-task  DVS  algorithm [16] (intraShin) and Gruian’s’ algorithm [5] (intraGruian) are used as representative algorithms of the path-based method and the stochastic method, respectively. (The details of these algorithms were described in Section 2.

\section{Results}

\section{Related Work}

@@

\bibitem{ahmed-chakrabarti}J.~Ahmed and C.~Chakrabarti, \emph{A dynamic task scheduling algorithm for battery powered DVS systems}, 2004 IEEE International Symposium on Circuits and Systems, Vancouver, BC, 2004. doi: 10.1109/ISCAS.2004.1329396

@@

@@

@@

@@

\bibitem{w-s-liu}W.-S.~Liu. \emph{Real-Time Systems}. Prentice Hall, Englewood Cliffs, NJ, June 2000.

@@

@@

\bibitem{yuan-nahrstedt}W.~Yuan and K.~Nahrstedt, \emph{Energy-efficient soft real-time CPU scheduling for mobile multimedia systems},  ACM SIGOPS Operating Systems Review, Volume 37, Pages 149-163, 2003

>>>>

\section{DVS Algorithms}
There are two kinds of voltage scheduling approaches for hard real-time systems depdending on the voltage scaling granularity: intra-task DVS (IntraDVS) and inter-task DVS (InterDVS). The intra-task DVS algorithms adjust the voltage within an individual task boundary, while the inter-task DVS algorithms determine the voltage on a task-by-task basis at each scheduling point. The main difference between them is whether the slack times are used for the current task or for the tasks that follow.  InterDVS algorithms distribute the slack times from the current task for the following tasks, while IntraDVS algorithms use the slack times from the current task for the current task itself.

@@

\section{Representative Algorithms} \label{hard-rt-algs}

%Table 2 summarizes the DVS algorithms selected for the comparative  study.

Nine  InterDVS  algorithms  are chosen, two of which are based on the RM scheduling policy, while the other seven algorithms are based on the EDF scheduling policy. Two algorithms were selected for IntraDVS, one from path-based IntraDVS algorithms, and the other from stochastic methods.

%In these selected DVS algorithms, one or sometimes more than one slack estimation methods explained in the previous section were used.

%In \texttt{lppsEDF} and \texttt{lppsRM} which were proposed by Shin \emph{et. al.} in  [18], slack time of a task is estimated using the maximum constant speed and Stretching-to-NTA methods. 
%
%The \texttt{ccRM} algorithm proposed by Pillai \emph{et. al.} [14]  is similar to \texttt{lppsRM} in the sense that it uses both the maximum constant speed and the Stretching-to-NTA methods. However, while \texttt{lppsRM} can adjust the voltage and clock speed only when a single task is active, \texttt{ccRM} extends the stretching to NTA method to the case where multiple tasks are active.

\begin{itemize}

    \item \texttt{lppsEDF} and \texttt{lppsRM} which were proposed in \cite{shin}, slack time of a task is estimated using the maximum constant speed and Stretching-to-NTA methods.

    \item \texttt{ccRM} algorithm proposed \cite{pillai} is similar to \texttt{lppsRM} in the sense that it uses both the maximum constant speed and the Stretching-to-NTA methods. However, while \texttt{lppsRM} can adjust the voltage and clock speed only when a single task is active, \texttt{ccRM} extends the stretching to NTA method to the case where multiple tasks are active.

    \item In \cite{pillai}, two other DVS algorithms are proposed, \texttt{ccEDF} and \texttt{laEDF}, for EDF scheduling policy. These algorithms estimate slack time of a task using the utilization updating method. While \texttt{ccEDF} adjusts the voltage and clock speed based on run-time variation in processor utilization alone, \texttt{laEDF} takes a more aggressive approach by estimating the amount of work required to be completed before NTA.

    \item \texttt{DRA} proposed in \cite{aydin}, is a representative DVS algorithm that are based on the priority-based slack stealing method. The algorithm estimates the slack time of a task using the priority-based slack stealing method along with the maximum constant speed and the Stretching-to-NTA methods.

    \item \texttt{AGR}, an extension of \texttt{DRA} proposed in \cite{aydin} is intended for more aggressive slack estimation and voltage/clock scaling. In \texttt{AGR}, in addition to the priority-based slack stealing, more slack times are identified by computing the amount of work required to be completed before NTA.

    \item \texttt{lpSEH} in \cite{kim} is based on the priority-based slack stealing method. It extends the priority-based slack stealing method by adding a procedure that estimates the slack time from lower-priority tasks that were completed earlier than expected. \texttt{DRA}, \texttt{AGR}, and \texttt{lpSEH} algorithms are somewhat similar to one another in the sense that all of them use the maximum constant speed in the off-line phase and the Stretching-to-NTA method in the on-line phase in addition to the priority-based slack stealing method.

    \item \texttt{darEDF} in \cite{zhuo-chakrabarti} is based on an efficient slack task scheduling scheme that employs dynamic speed setting slack in run queue. The algorithm uses a dynamic version of the average rate heuristic first and delays slack absorption, resulting in better battery performance.

\end{itemize}

Shin's intra-task DVS algorithm \cite{shin-kim-lee} and Gruian's algorithm \cite{gruian} are used as representative IntraDVS algorithms which use path-based method and the stochastic method, respectively.

%Pillai \emph{et. al.} also proposed two other DVS algorithms [14], \texttt{ccEDF} and \texttt{laEDF}, for EDF scheduling policy. These algorithms estimate slack time of a task using the utilization updating method. While \texttt{ccEDF} adjusts the voltage and clock speed based on run-time variation in processor utilization alone, laEDF takes a more aggressive approach by estimating the amount of work required to be completed before NTA.

%\texttt{DRA} and \texttt{AGR} , which were proposed by Aydin \emph{et. al.} in [2],  are two representative DVS algorithms that are based on the priority-based slack stealing method. The \texttt{DRA} algorithm estimates the slack time of a task using the priority-based slack stealing method along with the maximum constant speed and the Stretching-to-NTA methods.  Aydin \emph{et. al.} also extended the \texttt{DRA} algorithm and proposed another DVS algorithm called \texttt{AGR} for more aggressive slack estimation and voltage/clock scaling. In AGR, in addition to the priority-based slack stealing, more slack times are identified by computing the amount of work required to be completed before NTA. \texttt{lpSEH} is another  DVS  algorithm  which  is  based  on the priority-based slack stealing method [10]. Unlike \texttt{DRA} and \texttt{AGR}, \texttt{lpSEH} extends the priority-based slack stealing method by adding a procedure that estimates the slack time from lower-priority tasks that were completed earlier than expected. \texttt{DRA}, \texttt{AGR}, and \texttt{lpSEH} algorithms are somewhat similar to one another in the sense that all of them use the maximum constant speed in the off-line phase and the Stretching-to-NTA method in the on-line phase in addition to the priority-based slack stealing method. For IntraDVS algorithms, Shin's intra-task DVS  algorithm [16] (intraShin) and Gruian's' algorithm [5] (intraGruian) are used as representative algorithms of the path-based method and the stochastic method, respectively.
%
%\texttt{darEDF} is based on an efficient slack task scheduling scheme that employs dynamic speed setting slack in run queue. The algorithm uses a dynamic version of the average rate heuristic first and delays slack absorption, resulting in better battery performance.

\section{Results}
In \cite{kim-2} a performance comparison of most of the algorithms presented in section \ref{hard-rt-algs} was made using a simulation environment named SimDVS. It is argued that the energy efficiency of InterDVS algorithms depends significantly of slack estimation and the appropriateness of slack distribution. The following highlights the results obtained when evaluating InterDVS algorithms.

\paragraph{Impact of the number of task on the energy consumption} as the number of tasks increases,  the  energy  efficiency  of \texttt{lppsEDF}, \texttt{lppsRM}, and \texttt{ccRM} that only use the Stretching-to-NTA technique do not significantly improve, while that of the other more aggressive InterDVS algorithms improves significantly when compared with the energy consumption of an applicacion running on a DVS-unaware system with a power-down mode only. In the Stretching-to-NTA method, the slack time that can be exploited is limited to the time between the completion of a task instance and the arrival time of the next task instance, which is largely independent of the number of tasks in the system. For the other InterDVS algorithms, since the slack times can be taken from any completed task instance, as the number of task increases, each task has more slack sources and can be scheduled with a lowered clock speed.

\paragraph{Worst case processor utilization of task set} except for \texttt{lppsEDF}, the energy consumption of InterDVS algorithms increases as a linear function of WCPU of a task  set. For \texttt{lppsEDF}, the energy consumption increases faster than a linear function of WCPU of a task set. This indirectly indicates that the dynamic slack estimation method of \texttt{lppsEDF} is not very effective. 
\texttt{lppsEDF} shows better energy efficiency than \texttt{ccEDF} when WCPU is less than 0.7. In \texttt{ccEDF}, the clock speed is determined using the actual processor utilization at the scheduling point.  Since the actual processor utilization increases when a low-speed task instance completes its execution, the next task instance needs to be executed in a higher speed. Such  voltage fluctuation occurs more often as the WCPU decreases. Thus, as the WCPU decreases, the energy efficiency of ccEDF becomes worse than that of \texttt{lppsEDF}. The results for \texttt{lppsRM} and \texttt{ccRM} are very similar to that of \texttt{lppsEDF}.

\paragraph{Speed bound} the energy efficiency of \texttt{AGR} and \texttt{lpSEH} is very close to the theoretical lower bound\footnote{Determined by the Yao's algorithm in \cite{yao}} when the speed bound factor is near 0.5. For the aggressive InterDVS algorithms, the energy efficiency is highest when the speed bound factor was set to average case processor utilization (ACPU). When the selected speed bound factor is close to ACPU $(= 0,55 \times \textnormal{WCPU})$, the best energy effciency is achieved for \texttt{laEDF}. For \texttt{ccEDF} this trend does not hold. For RM InterDVS algorithms, the performance gap between the energy efficiency and that of the theoretical lower bound was roughly 35$\sim$40\%.

\textbf{==> Aqui Poner resultados de \texttt{darEDF} <==}

\subsection{Performance evaluation of Intra-Task DVS algorithms}
The two representative IntraDVS algorithms perform quite differently depending on available slack times. When the relative energy consumption ratio of intraGruian over intraShin was measured, if the ratio is larger than 1, intraGruian performs better than intraShin. When  the  slack  ratio  is  less  than  1.2, intraShin outperforms intraGruian because intraShin spends more time in the lower speed region than intraGruian. When the slack ratio is increased, intraGruian spends more time in the lower speed region than intraShin.

\section{Related Work}
\begin{itemize}
    \item otros algoritmos para hard-time (lfpsEDF por ejemplo)
    \item scheduling en mobile devices basado en offloading
\end{itemize}

@@

@@

\bibitem{aydin}H.~Aydin, R.~Melhem, D.~Mosse, and P.M.~Alvarez. \emph{Dynamic  and  Aggressive  Scheduling  Techniques  for  Power-Aware  Real-Time Systems}. In Proceedings  of IEEE Real-Time Systems Symposium, December 2001.

@@

\bibitem{gruian}F.~Gruian. \emph{Hard Real-Time Scheduling Using Stochastic Data and  DVS  Processors}. In Proceedings  of  the  International Symposium on Low Power Electronics and Design , pages 46–51, August 2001.

@@

\bibitem{kim}W.~Kim, J.~Kim, and S.L.~Min. \emph{A Dynamic Voltage Scaling  Algorithm  for  Dynamic-Priority  Hard  Real-Time  Systems Using Slack Time Analysis}. In Proceedings of Design, Automation and Test in Europe (DATE'02), pages 788–794, March 2002.

\bibitem{kim-2}W.~Kim, J.~Kim, and S.L.~Min \emph{Performance comparison of dynamic voltage scaling algorithms for hard real-time systems}. proc. RTAS. pp 219-228. 2002.

@@

\bibitem{w-s-liu}W.S.~Liu. \emph{Real-Time Systems}. Prentice Hall, Englewood Cliffs, NJ, June 2000.

@@

\bibitem{pillai}P.~ Pillai and K.G.~Shin. \emph{Real-Time Dynamic Voltage Scaling
for Low-Power Embedded Operating Systems}.  In Proceedings of 18th ACM Symposium on Operating Systems Principles (SOSP’01), pages 89–102, October 2001.

@@

\bibitem{shin-kim-lee}D.~Shin, J.~Kim, and S.~Lee.  \emph{Intra-Task Voltage Scheduling for Low-Energy Hard Real-Time Applications}. IEEE Design and Test of Computers, 18(2):20–30, March 2001

\bibitem{shin}Y.~Shin, K.~Choi, and  T.~Sakurai. \emph{Power  Optimization  of Real-Time  Embedded  Systems  on  Variable  Speed  Processors}. In Proceedings of the International Conference on Computer-Aided Design, pages 365–368, November 2000.

\bibitem{yao}F.~Yao, A.~Demers, and A.~Shenker. \emph{A Scheduling Model for Reduced CPU Energy}. In Proceedings of the IEEE Foundations of Computer Science, pages 374–382, October 1995.

\bibitem{yuan-nahrstedt}W.~Yuan and K.~Nahrstedt, \emph{Energy-efficient soft real-time CPU scheduling for mobile multimedia systems},  ACM SIGOPS Operating Systems Review, Volume 37, Pages 149-163, 2003.

\bibitem{zhuo-chakrabarti}J.~Zhuo and C.~Chakrabarti, \emph{An efficient dynamic task scheduling algorithm for battery powered DVS systems}, 2004 IEEE International Symposium on Circuits and Systems, Vancouver, BC, 2004. doi: 10.1109/ISCAS.2004.1329396
